Near Optimal Behavior with Approximate Abstraction

### 2016-02-04 paper/introduction.tex
Some intro commit

\section{Introduction}
Abstraction plays a fundamental role in learning. Through abstraction, intelligent agents may reason about only the salient features of their environment while ignoring what is irrelevant. Consequently, agents are able to solve considerably more complex problems than they would be able to without the use of abstraction. However, \textit{exact abstractions}, which treat fully-identical situations alike, require complete knowledge that is often computationally intractable to obtain. Furthermore, often no two situations are identical, so exact abstraction is ineffective. To overcome these issues we investigate \textit{approximate abstractions} that enable agents to treat sufficiently similar situations as identical. This work characterizes the impact of equating ``sufficiently similar'' states in the context of planning and \ac{RL} in \acp{MDP}. The remainder of our introduction contextualizes these intuitions in \acp{MDP}.

% Abstraction is a good idea b/c MDP has poly comp comp and samp comp in S, but S grows fast.
Solving for optimal behavior in \acp{MDP} in a planning setting is known to be P-Complete in the size of the state space~\cite{papadimitriou1987complexity,littman1995complexity}. Similarly, many \ac{RL} algorithms for solving \acp{MDP} are known to require a number of samples polynomial in the size of the state space~\cite{strehl2009reinforcement}. Although polynomial runtime or sample complexity may seem like a reasonable constraint, due to Bellman's curse of dimensionality, the size of the state space of an \ac{MDP} grows super-polynomially with the number of variables that characterize the domain. Thus, bounds polynomial in state space size often translate to intractable problems for sufficiently complicated tasks. \dnote{Would prefer a different example} For instance, a robot involved in a pick and place task might be able to employ planning algorithms to solve for how to manipulate some objects into a desired configuration in time polynomial in the number of states, but the number of states it must consider to do so grows exponentially with the number of objects it plans over \enote{Cite Dabe Icaps}.

% Existing work on state abstraction.
Thus, a key research agenda for planning and \ac{RL} is leveraging abstraction to reduce large state spaces~\cite{andre2002state,jong2005state,dietterich2000hierarchical,Bean2011}. These methods are characterized by reducing \textit{ground} \acp{MDP} with large state spaces to \textit{abstract} MDPs with smaller state spaces by aggregating states according to some notion of equality or similarity. In the context of \acp{MDP}, we understand exact abstractions as those that aggregate states with equal values of particular quantities, e.g. optimal $Q$-values. Existing work has characterized how exact abstractions can fully maintain optimality~\cite{li2006towards,dean1997modelmin}.

% Thesis: Approximate abstraction for three reasons.
However, performing approximate abstraction in \acp{MDP} by relaxing the state aggregation criteria from equality of quantities to $\epsilon$-closeness offers three benefits. First, approximate abstractions employ the sort of knowledge that we expect a planning or learning algorithm to compute without fully solving the \ac{MDP}. In contrast, exact abstractions often require solving for optimal behavior, thereby defeating the purpose of abstraction. Second, because of their relaxed criteria, approximate abstractions can achieve greater degrees of compression than exact abstractions. This is particularly important in environments where no two states are identical. \dnote{note this in empirical results discussion, that random is sort of like a natural environment and when eps is 0, no compression} Third, because the state aggregation criteria are relaxed to approximate equality, they are able to tune the aggressiveness of abstraction by adjusting what they consider sufficiently similar states. The thesis of this work is that aggregation of states on the basis of various approximate criteria can reduce state space size while incurring only bounded error in the resulting behavior.

>>>>

\section{Introduction}
Abstraction plays a fundamental role in learning. Through abstraction, intelligent agents may reason about only the salient features of their environment while ignoring what is irrelevant. Consequently, agents are able to solve considerably more complex problems than they would be able to without the use of abstraction. However, \textit{exact abstractions}, which treat fully-identical situations alike, require complete knowledge that is often computationally intractable to obtain. Furthermore, often no two situations are identical, so exact abstractions are often ineffective. To overcome these issues we investigate \textit{approximate abstractions} that enable agents to treat sufficiently similar situations as identical. This work characterizes the impact of equating ``sufficiently similar'' states in the context of planning and \ac{RL} in \acp{MDP}. The remainder of our introduction contextualizes these intuitions in \acp{MDP}.

% Abstraction is a good idea b/c MDP has poly comp comp and samp comp in S, but S grows fast.
Solving for optimal behavior in \acp{MDP} in a planning setting is known to be P-Complete in the size of the state space~\cite{papadimitriou1987complexity,littman1995complexity}. Similarly, many \ac{RL} algorithms for solving \acp{MDP} are known to require a number of samples polynomial in the size of the state space~\cite{strehl2009reinforcement}. Although polynomial runtime or sample complexity may seem like a reasonable constraint, the size of the state space of an \ac{MDP} grows super-polynomially with the number of variables that characterize the domain, a results of Bellman's curse of dimensionality. Thus, solutions polynomial in state space size are often ineffective for sufficiently complex tasks. \dnote{Would prefer a different example} For instance, a robot involved in a pick and place task might be able to employ planning algorithms to solve for how to manipulate some objects into a desired configuration in time polynomial in the number of states, but the number of states it must consider to do so grows exponentially with the number of objects it plans over \enote{Cite Dabe Icaps}.

% Existing work on state abstraction.
Thus, a key research agenda for planning and \ac{RL} is leveraging abstraction to reduce large state spaces~\cite{andre2002state,jong2005state,dietterich2000hierarchical,Bean2011}. This agenda has given rise to methods that reduce \textit{ground} \acp{MDP} with large state spaces to \textit{abstract} MDPs with smaller state spaces by aggregating states according to some notion of equality or similarity. In the context of \acp{MDP}, we understand exact abstractions as those that aggregate states with equal values of particular quantities, e.g. optimal $Q$-values. Existing work has characterized how exact abstractions can fully maintain optimality in \acp{MDP}~\cite{li2006towards,dean1997modelmin}.

% Thesis: Approximate abstraction for three reasons.
Performing approximate abstraction in \acp{MDP} by relaxing the state aggregation criteria from equality of quantities to $\epsilon$-closeness offers three benefits. First, approximate abstractions employ the sort of knowledge that we expect a planning or learning algorithm to compute without fully solving the \ac{MDP}. In contrast, exact abstractions often require solving for optimal behavior, thereby defeating the purpose of abstraction. Second, because of their relaxed criteria, approximate abstractions can achieve greater degrees of compression than exact abstractions. This difference is particularly important in environments where no two states are identical. \dnote{note this in empirical results discussion, that random is sort of like a natural environment and when eps is 0, no compression} Third, because the state aggregation criteria are relaxed to approximate equality, approximate abstractions are able to tune the aggressiveness of abstraction by adjusting what they consider sufficiently similar states. The thesis of this work is that aggregation of states on the basis of various approximate criteria can reduce state space size while incurring only bounded error in the resulting behavior.
